---
title: "finalproject2"
author: "Andro Manukov"
date: "May 10, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Installing Packages:
```{r Installing Packages}
library(tidyverse)
library(GGally)
library(ggplot2)
library(scales)
library(gridExtra)
library(grid)
library(randomForest)
library(lmtest)
```

Reading csv and subsetting to just city hotels:
```{r Reading Data and Subsetting}
Hotel <- read_csv("D:/UIUC/Spring2020/stat425/finalproject/stat425_fpdata.csv")
city_hotels <- subset(Hotel, Hotel$hotel == "City Hotel") #Subset just city hotels because im section two. 
```

Creating seasons variable:
```{r Creating Seasons}
city_hotels$arrival_season = "" #init variable
i = 1 #index for rows
for(val in city_hotels$arrival_date_month) #increment through each row
{
  if(val == "January" | val == "February" | val == "December") { #conditional
  city_hotels$arrival_season[i] = "Winter" #change to season
} else if(val == "March" | val == "May" | val == "April") {
  city_hotels$arrival_season[i] = "Spring"
} else if(val == "June" | val == "July" | val == "August") {
  city_hotels$arrival_season[i] = "Summer"
} else {
  city_hotels$arrival_season[i] = "Fall"
}
  
i = i + 1 #incremenet index
}

```

```{r}
head(city_hotels)
```


Creating total stays variable:
```{r}
city_hotels$total_stays <- city_hotels$stays_in_weekend_nights + city_hotels$stays_in_week_nights #creating total stays variable
```

Creating Bar Plots:
```{r}
#READ THIS ONE FOR A GENERAL TREND OF ALL OF THE OTHER BAR PLOTS!!!!!!!!!!!!
#THIS BAR PLOT IS FOR ADR SEPERATED BY YEAR!!!!!!!!!!

year_group <- summarise_at(group_by(city_hotels, arrival_date_year), vars(adr), funs(mean(.,na.rm=FALSE))) #GROUPBY THE VARIABLE WE ARE INVESTIGATING

ggplot(data = year_group, aes(x = arrival_date_year, y = adr, fill = arrival_date_year)) + 
  geom_bar(position = 'dodge', stat='identity') + #create bar plot
  geom_text(aes(label= dollar(adr)), position=position_dodge(width=0.9), vjust=-0.25) + #add text to bar plot
  xlab("Year") + ylab("Average Daily Rate") + ggtitle("Average Daily Rate seperated by Year") + scale_y_continuous(labels = scales::dollar) + theme(legend.position = "none", panel.grid =   element_blank(), #add all titles and everything
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) #ADD X, Y, TITLE LABELS AND ADJUST THEM
```

```{r}
#plot for Average Daily Rate seperated by Season
season_group <- summarise_at(group_by(city_hotels, arrival_season), vars(adr), funs(mean(.,na.rm=FALSE)))

ggplot(data = season_group, aes(x = arrival_season, y = adr, fill = arrival_season)) +
  geom_bar(position = 'dodge', stat='identity') +
  geom_text(aes(label= dollar(adr)), position=position_dodge(width=0.9), vjust=-0.25) +
  xlab("Season") + ylab("Average Daily Rate") + ggtitle("Average Daily Rate seperated by Season") +   scale_y_continuous(labels = scales::dollar) + theme(legend.position = "none", panel.grid =   element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

```{r}
#plot for Average Daily Rate seperated by Arrival Date of Month"
day_group <- summarise_at(group_by(city_hotels, arrival_date_day_of_month), vars(adr), funs(mean(.,na.rm=FALSE)))

ggplot(data = day_group, aes(x = arrival_date_day_of_month, y = adr, fill = "red")) +
  geom_bar(position = 'dodge', stat='identity') +
  xlab("Day") + ylab("Average Daily Rate") + ggtitle("Average Daily Rate seperated by Arrival Date of Month") +   scale_y_continuous(labels = scales::dollar) + theme(legend.position = "none", panel.grid =   element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

```{r}
#creating plot to test for interaction between year and season
year_season_group <- summarise_at(group_by(city_hotels, arrival_season, arrival_date_year), vars(adr), funs(mean(.,na.rm=FALSE)))

ggplot(year_season_group, aes(x=arrival_season, y=adr, fill = as.factor(arrival_date_year))) +
    geom_bar(position = 'dodge', stat='identity') +
  xlab("Season") + ylab("Average Daily Rate") + ggtitle("Testing for Interaction between Year and Season") +   scale_y_continuous(labels = scales::dollar) + theme(panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) + labs(fill = "Year")
```

```{r}
#creating is canceled group plot.
is_canceled_group <- summarise_at(group_by(city_hotels, is_canceled), vars(adr), funs(mean(.,na.rm=FALSE)))

ggplot(data = is_canceled_group, aes(x = is_canceled, y = adr, fill = is_canceled)) +
  geom_bar(position = 'dodge', stat='identity') +
  geom_text(aes(label= dollar(adr)), position=position_dodge(width=0.9), vjust=-0.25) +
  xlab("is_canceled") + ylab("Average Daily Rate") + ggtitle("Average Daily Rate seperated by is_canceled") +   scale_y_continuous(labels = scales::dollar) + theme(legend.position = "none", panel.grid =   element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

```{r}
#creating plot for meal group
meal_group <- summarise_at(group_by(city_hotels, meal), vars(adr), funs(mean(.,na.rm=FALSE)))

meal_package_plot <- ggplot(data = meal_group, aes(x = meal, y = adr, fill = meal)) +
  geom_bar(position = 'dodge', stat='identity') +
  geom_text(aes(label= dollar(adr)), position=position_dodge(width=0.9), vjust=-0.25) +
  xlab("Meal Package") + ylab("Average Daily Rate") + ggtitle("") +   scale_y_continuous(labels = scales::dollar) + theme(legend.position = "none", panel.grid =   element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

```{r}
#mcreating plot for market_group
market_group <- summarise_at(group_by(city_hotels, market_segment), vars(adr), funs(mean(.,na.rm=FALSE)))

market_segment_plot <- ggplot(data = market_group, aes(x = market_segment, y = adr, fill = market_segment)) +
  geom_bar(position = 'dodge', stat='identity') +
  geom_text(aes(label= dollar(adr), angle = 90), hjust= 2) +
  xlab("Market Segment") + ylab("Average Daily Rate") + ggtitle("Average Daily Rate seperated by Market Segment and Seperated by Meal Plan") +   scale_y_continuous(labels = scales::dollar) + theme(legend.position = "none", panel.grid =   element_blank(),
        plot.title = element_text(hjust = 0),
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))
```
 
```{r}
grid.arrange(meal_package_plot, market_segment_plot, layout_matrix = cbind(c(2,2), c(1,1))) #combining to save space
```

```{r}
#creating room type plot
reserve_group <- summarise_at(group_by(city_hotels, reserved_room_type), vars(adr), funs(mean(.,na.rm=FALSE)))

reserve_group_plot <- ggplot(data = reserve_group, aes(x = reserved_room_type, y = adr, fill = reserved_room_type)) +
  geom_bar(position = 'dodge', stat='identity') +
  geom_text(aes(label= dollar(adr), angle = 90), hjust = 2) +
  xlab("Room Type") + ylab("Average Daily Rate") + ggtitle("") +   scale_y_continuous(labels = scales::dollar) + theme(legend.position = "none", panel.grid =   element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

```{r}
#creating customer group plot
customer_group <- summarise_at(group_by(city_hotels, customer_type), vars(adr), funs(mean(.,na.rm=FALSE)))

customer_group_plot <- ggplot(data = customer_group, aes(x = customer_type, y = adr, fill = customer_type)) +
  geom_bar(position = 'dodge', stat='identity') +
  geom_text(aes(label= dollar(adr), angle = 90), hjust = 3) +
  xlab("Customer Type") + ylab("Average Daily Rate") + ggtitle("Average Daily Rate seperated by Customer Type and Seperated by Room Type") +   scale_y_continuous(labels = scales::dollar) + theme(legend.position = "none", panel.grid =   element_blank(),
        plot.title = element_text(hjust = 0),
        plot.subtitle = element_text(hjust = 0.5))
```

```{r}
#combining two plots to save space :(
grid.arrange(reserve_group_plot, customer_group_plot, layout_matrix = cbind(c(2,2), c(1,1)))
```

```{r}
#looking at Average Daily Rate seperated by Number of Special Requests
special_group <- summarise_at(group_by(city_hotels, total_of_special_requests), vars(adr), funs(mean(.,na.rm=FALSE)))

ggplot(data = special_group, aes(x = total_of_special_requests, y = adr, fill = "red")) +
  geom_bar(position = 'dodge', stat='identity') +
  geom_text(aes(label= dollar(adr)), position=position_dodge(width=0.9), vjust=-0.25) +
  xlab("Number of Special Requests") + ylab("Average Daily Rate") + ggtitle("Average Daily Rate seperated by Number of Special Requests") +   scale_y_continuous(labels = scales::dollar) + theme(legend.position = "none", panel.grid =   element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

```{r}
#looking at Average Daily Rate seperated by Number of Total People"
city_hotels$total_people <- city_hotels$adults + city_hotels$children

people_group <- summarise_at(group_by(city_hotels, total_people), vars(adr), funs(mean(.,na.rm=FALSE)))

ggplot(data = people_group, aes(x = total_people, y = adr, fill = total_people)) +
  geom_bar(position = 'dodge', stat='identity') +
  geom_text(aes(label= dollar(adr)), position=position_dodge(width=0.9), vjust=-0.25) +
  xlab("Total People") + ylab("Average Daily Rate") + ggtitle("Average Daily Rate seperated by Number of Total People") +   scale_y_continuous(labels = scales::dollar) + theme(legend.position = "none", panel.grid =   element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

```{r}
head(city_hotels)
```

Testing Numerical Variables:
```{r}
ggpairs(select(city_hotels, c(adr, total_stays, lead_time))) + ggtitle("Analysis between numerical total_stays and lead_time and ADR") + theme(legend.position = "none", panel.grid =   element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
#looking at total stays and lead time
```


Subsetting Dataset once more:
```{r Modifying Dataset}
city_hotels_removed <- select(city_hotels, is_canceled, reserved_room_type, arrival_season, total_people, meal, market_segment, customer_type, arrival_date_year, adr) #subsetting data
```

Converting to factor variables:
```{r}
city_hotels_removed$is_canceled = as.factor(city_hotels$is_canceled)
city_hotels_removed$reserved_room_type = as.factor(city_hotels$reserved_room_type)
city_hotels_removed$arrival_season = as.factor(city_hotels$arrival_season)
city_hotels_removed$total_people = as.factor(city_hotels$total_people)
city_hotels_removed$meal = as.factor(city_hotels$meal)
city_hotels_removed$market_segment = as.factor(city_hotels$market_segment)
city_hotels_removed$customer_type = as.factor(city_hotels$customer_type)
city_hotels_removed$arrival_date_year = as.factor(city_hotels$arrival_date_year)
#creating factor variables
```

```{r}
head(city_hotels_removed)
```

Creating simple model:
```{r}
simple_model <- lm(adr ~ . , data = city_hotels_removed)
summary(simple_model) #fitting simple model, looking at summary
```

BIC of simple model:
```{r}
BIC(simple_model) #BIC for simple model
```



ADD A TRAIN AND TEST SET FOR 3.2 AND 3.3

Subsetting data to remove marketsegment:
```{r}
city_hotels_removed <- select(city_hotels_removed, is_canceled, reserved_room_type, arrival_season, total_people, meal, customer_type, arrival_date_year, adr) #cutting down dataset
```

Creating test and train sets:
```{r}
# creating test and train datasets
## 75% of the sample size
smp_size <- floor(0.75 * nrow(city_hotels_removed))

## set the seed to make your partition reproducible
set.seed(123)
train_ind <- sample(seq_len(nrow(city_hotels_removed)), size = smp_size) #index

train <- city_hotels_removed[train_ind, ]
test <- city_hotels_removed[-train_ind, ]

#https://stackoverflow.com/questions/17200114/how-to-split-data-into-training-testing-sets-using-sample-function
```



Creating complex model:
```{r}
complex_model <- lm(adr ~ . + arrival_season * arrival_date_year, data = train)
summary(complex_model) #fitting model and looking at summary
```

BIC of complex model:
```{r}
BIC(complex_model) #BIC of MLR model
```

Predict using complex model:
```{r}
#predict_df = data.frame(is_canceled = "0", arrival_season = "Winter", arrival_date_year = #"2016", meal = "BB", reserved_room_type = "C", customer_type = "Transient", total_people #= "4")

mean(predict(complex_model, test)) #predicting the ADR from the MLR model using test data
```

```{r}
head(city_hotels_removed) #viewing dataset for convience
```

Creating Random Forest:
```{r}
rfModel = randomForest(adr ~ ., ntree = 2000, data = train, seed = 425)
rfModel #creating the RF
```

Importance values of rf:
```{r}
rfModel$importance #importance of RF model variables
```

RMSE of RF:
```{r}
mean(sqrt(rfModel$mse)) #rmse of RF model
```

Predict using RF:
```{r}
predict_tree = predict(rfModel, data = test, predict.all = TRUE)
mean(predict_tree) #predict using RF
``` 

RMSE of MLR:
```{r}
sqrt(sum(complex_model$residuals^2)/nrow(city_hotels_removed)) #RMSE of MLR model
```

SD of ADR:

```{r}
sd(test$adr)
```

Model Diagnostics:

QQ plot
```{r}
qqnorm(complex_model$residuals) #qq plot to test for normality 
```
Residual vs Fitted
```{r}
#plot for testing for constance variance assumption
plot(complex_model$fitted.values, complex_model$residuals, xlab = "Fitted", ylab = "Residuals", main = "Residuals vs Fitted")
```
Residual vs index
```{r}
#plot to help test for correlated errors
len = 1:1213 #index
plot(len, complex_model$residuals, xlab = "Index", ylab = "Residuals", main = "Residuals vs Index")
lines(len, complex_model$residuals)
```

durbin watson test
```{r}
dwtest(complex_model) #correlated errors test
```

